---
title: "Project 2"
subtitle: "Descriptive Statistics"
author: "Denys, Zakhar, Kvitoslava"
institute: Kyiv School of Economics
date: "today"
format:
  html:
    toc: true
    toc-depth: 3
execute:
  engine: knitr
---

# Packages

```{r}
library(tidyverse)
# library(tidymodels)
# library(DAAG)
```

# Data

Here are the links to the data files that we will be using in this project:

- [2022 Russia Ukraine War](https://www.kaggle.com/datasets/piterfm/2022-ukraine-russian-war) (updated weekly)

- [2022 Russia Ukraine War, Losses, Oryx + Images](https://www.kaggle.com/datasets/piterfm/2022-ukraine-russia-war-equipment-losses-oryx)

- [Massive Missile Attacks on Ukraine](https://www.kaggle.com/datasets/piterfm/massive-missile-attacks-on-ukraine?select=missiles_and_uav.csv)

```{r}
general_folder <- "Project 2/data"
ru_ua_folder <- "2022 Russia Ukraine War"
ru_ua_oryx_folder <- "2022 Russia Ukraine War, Losses, Oryx + Images"
missile_atcks_folder <- "Massive Missile Attacks on Ukraine"

ru_loss_persons <- read.csv(file.path(general_folder, ru_ua_folder, "russia_losses_personnel.csv"))
ru_loss_equipment <- read.csv(file.path(general_folder, ru_ua_folder, "russia_losses_equipment.csv"))
ru_loss_equipment_correction <- read.csv(file.path(general_folder, ru_ua_folder, "russia_losses_equipment_correction.csv"))

losses_ru <- read.csv(file.path(general_folder, ru_ua_oryx_folder, "losses_russia.csv"))
losses_ua <- read.csv(file.path(general_folder, ru_ua_oryx_folder, "losses_ukraine.csv"))

missiles_daily <- read.csv(file.path(general_folder, missile_atcks_folder, "missile_attacks_daily.csv"))
missiles_and_uav <- read.csv(file.path(general_folder, missile_atcks_folder, "missiles_and_uav.csv"))
```

# Data Overview

...

# Data Processing

## Equipment Losses

### Russia

First, we will deal with the Russia equipment losses correction!

```{r}
apply_corrections <- function(main_data, corrections) {
    corrected_data <- main_data %>%
        left_join(corrections, by = "date", suffix = c("", "_correction"))
        
    corrected_columns <- c(
      "aircraft", "helicopter", "tank", "APC", "field_artillery",
      "MRL", "drone", "naval_ship", "submarines", 
      "anti_aircraft_warfare", "special_equipment", 
      "vehicles_and_fuel_tanks", "cruise_missiles"
    )
    
    for (col in corrected_columns) {
        correction_col <- paste0(col, "_correction")
        if (correction_col %in% colnames(corrected_data)) {
            # replace NA in correction with 0 (на всяк випадок)
            corrected_data[[correction_col]][is.na(corrected_data[[correction_col]])] <- 0
            
            # cumulative correction
            corrected_data[[col]] <- corrected_data[[col]] + cumsum(replace_na(corrected_data[[correction_col]], 0))
        }
    }
    
    # Remove correction columns
    corrected_data <- corrected_data %>%
        select(-ends_with("_correction"))
    
    return(corrected_data)
}
```

```{r}
# make both datasets have the same ascending date order

ru_loss_equipment <- ru_loss_equipment %>%
    arrange(date)

# ru_loss_equipment %>% 
#     head() %>%
#     select(date, day)

ru_loss_equipment_correction <- ru_loss_equipment_correction %>%
    arrange(date)

ru_loss_equipment_corrected <- apply_corrections(ru_loss_equipment, ru_loss_equipment_correction)

ru_loss_equipment_corrected <- ru_loss_equipment_corrected %>% 
    arrange(desc(date))

write.csv(ru_loss_equipment_corrected, file.path(general_folder, ru_ua_folder, "russia_losses_equipment_corrected.csv"), row.names = FALSE)
```

Now, we will evaluate the score for each type of equipment!

```{r}
weights <- list(
  aircraft = 69,                 # Aircraft: High cost and strategic value
  helicopter = 50,               # Helicopter: Tactical support and mobility
  tank = 30,                     # Tank: Backbone of ground offensives
  APC = 10,                      # APC: Troop transport
  field.artillery = 15,          # Field Artillery: Combat support
  MRL = 40,                      # MRL: High-impact artillery strikes
  military.auto = 10,            # Military Auto: Logistics and mobility
  fuel.tank = 10,                # Fuel Tank: Operations sustainment
  drone = 20,                    # Drone: Reconnaissance and targeted strikes
  naval.ship = 100,              # Naval Ship: Maritime operations
  anti.aircraft.warfare = 60,    # Anti-Aircraft Warfare: Air defense
  special.equipment = 40,        # Special Equipment: Specialized operational roles
  mobile.SRBM.system = 100,      # Mobile SRBM System: Strategic deterrence
  vehicles.and.fuel.tanks = 30,  # Vehicles and Fuel Tanks: Logistics
  cruise.missiles = 40,          # Cruise Missiles: High-impact strikes
  submarines = 1000,             # Submarines: Strategic naval supremacy
  greatest.losses.direction = 1  # skip this column
)
```

```{r}
# rearange columns
ru_loss_equipment_temp <- ru_loss_equipment_corrected %>%
  select(-c(date, day, greatest.losses.direction))

# use weights to evaluate the score for each type of equipment
for (equipment in names(weights)) {
  if (equipment %in% colnames(ru_loss_equipment_temp)) {
    ru_loss_equipment_temp[[equipment]] <- ru_loss_equipment_temp[[equipment]] * weights[[equipment]]
  }
}

ru_loss_equipment_temp %>% 
  head()

# calculate the score for each row (cumulative)
ru_loss_equipment_temp$cdf_loss_score <- rowSums(ru_loss_equipment_temp, na.rm = TRUE)

ru_loss_equipment_evaluated <- cbind(ru_loss_equipment_corrected, cdf_loss_score = ru_loss_equipment_temp$cdf_loss_score)
glimpse(ru_loss_equipment_evaluated)
summary(ru_loss_equipment_temp$cdf_loss_score)
```

Now, we will evaluate the PDF for the equipment losses!

```{r}
ru_loss_equipment_evaluated <- ru_loss_equipment_evaluated %>% 
  arrange(date)
```

```{r}
ru_loss_equipment_evaluated_pdf <- ru_loss_equipment_evaluated %>%
  arrange(desc(date)) %>%
  mutate(pdf_loss_score = c(ru_loss_equipment_evaluated$cdf_loss_score[1], diff(ru_loss_equipment_evaluated$cdf_loss_score)))

ru_loss_equipment_evaluated_pdf %>% 
  head()
```

## Distribution of Equipment Losses

First, let's split the data by seasons (winter, spring, summer, autumn) + year

```{r}
SUMMER_HEX <- "#e31a1c"
WINTER_HEX <- "#1f78b4"
SPRING_HEX <- "#33a02c"
AUTUMN_HEX <- "#ff7f00"

disrtibution_data <- ru_loss_equipment_evaluated_pdf %>% 
  mutate(
    season = case_when(
      month(date) %in% c(12, 1, 2) ~ "Winter",
      month(date) %in% c(3, 4, 5) ~ "Spring",
      month(date) %in% c(6, 7, 8) ~ "Summer",
      month(date) %in% c(9, 10, 11) ~ "Autumn"
    ),
    year = year(date)
  ) %>%
  select(pdf_loss_score, season, year)
```

Now, let's build the distribution of equipment losses! by each season

```{r}
disrtibution_data %>%
  filter(year == 2022) %>%
  ggplot(
    aes(x = pdf_loss_score, fill = season)
    ) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c(
    "Winter" = WINTER_HEX,
    "Spring" = SPRING_HEX,
    "Summer" = SUMMER_HEX,
    "Autumn" = AUTUMN_HEX
  )) +
  labs(
    title = "Distribution of Equipment Losses in 2022",
    x = "PDF Loss Score",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )

```

```{r}
disrtibution_data %>%
  filter(year == 2023) %>%
  ggplot(
    aes(x = pdf_loss_score, fill = season)
    ) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c(
    "Winter" = WINTER_HEX,
    "Spring" = SPRING_HEX,
    "Summer" = SUMMER_HEX,
    "Autumn" = AUTUMN_HEX
  )) +
  labs(
    title = "Distribution of Equipment Losses in 2023",
    x = "PDF Loss Score",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )

```

```{r}
disrtibution_data %>%
  filter(year == 2024) %>%
  ggplot(
    aes(x = pdf_loss_score, fill = season)
    ) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c(
    "Winter" = WINTER_HEX,
    "Spring" = SPRING_HEX,
    "Summer" = SUMMER_HEX,
    "Autumn" = AUTUMN_HEX
  )) +
  labs(
    title = "Distribution of Equipment Losses in 2024",
    x = "PDF Loss Score",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  ) +
  scale_x_continuous(limits = quantile(disrtibution_data$pdf_loss_score, c(0, 0.95), na.rm = TRUE))
```

## Build two distributions for the hypothesis testing

  ```{r}
  disrtibution_data %>%
    filter((year == 2023 & season == "Summer") | (year == 2024 & season == "Autumn")) %>%
    ggplot(aes(x = pdf_loss_score, fill = season)) +
    geom_density(alpha = 0.5) +
    scale_fill_manual(values = c("Summer" = SUMMER_HEX, "Autumn" = AUTUMN_HEX)) +
    labs(
      title = "Distribution of Equipment Losses in Summer 2023 and Autumn 2024",
      x = "PDF Loss Score",
      y = "Density"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_x_continuous(limits = quantile(disrtibution_data$pdf_loss_score, c(0, 0.95), na.rm = TRUE))
```
