---
title: "Project 2"
subtitle: "Descriptive Statistics"
author: "Denys, Zakhar, Kvitoslava"
institute: Kyiv School of Economics
date: "today"
format:
  html:
    toc: true
    toc-depth: 3
execute:
  engine: knitr
---

# Packages

```{r}
library(tidyverse)
# library(tidymodels)
# library(DAAG)
```

# Data

Here are the links to the data files that we will be using in this project:

- [2022 Russia Ukraine War](https://www.kaggle.com/datasets/piterfm/2022-ukraine-russian-war) (updated weekly)

- [2022 Russia Ukraine War, Losses, Oryx + Images](https://www.kaggle.com/datasets/piterfm/2022-ukraine-russia-war-equipment-losses-oryx)

- [Massive Missile Attacks on Ukraine](https://www.kaggle.com/datasets/piterfm/massive-missile-attacks-on-ukraine?select=missiles_and_uav.csv)

```{r}
general_folder <- "Project 2/data"
ru_ua_folder <- "2022 Russia Ukraine War"
ru_ua_oryx_folder <- "2022 Russia Ukraine War, Losses, Oryx + Images"
missile_atcks_folder <- "Massive Missile Attacks on Ukraine"

ru_loss_persons <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/2022 Russia Ukraine War/russia_losses_personnel.csv")
ru_loss_equipment <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/2022 Russia Ukraine War/russia_losses_equipment.csv")
ru_loss_equipment_correction <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/2022 Russia Ukraine War/russia_losses_equipment_correction.csv")

losses_ru <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/2022 Russia Ukraine War, Losses, Oryx + Images/losses_russia.csv")
losses_ua <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/2022 Russia Ukraine War, Losses, Oryx + Images/losses_ukraine.csv")

missiles_daily <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/Massive Missile Attacks on Ukraine/missile_attacks_daily.csv")
missiles_and_uav <- read.csv("~/Documents/GitHub/scam-project/Project 2/data/Massive Missile Attacks on Ukraine/missiles_and_uav.csv")
```

# Data Overview

...

# Data Processing

## Equipment Losses

### Russia

First, we will deal with the Russia equipment losses correction!

```{r}
apply_corrections <- function(main_data, corrections) {
    corrected_data <- main_data %>%
        left_join(corrections, by = "date", suffix = c("", "_correction"))

    corrected_columns <- c("aircraft", "helicopter", "tank", "APC", "field_artillery",
                                                 "MRL", "drone", "naval_ship", "submarines",
                                                 "anti_aircraft_warfare", "special_equipment",
                                                 "vehicles_and_fuel_tanks", "cruise_missiles")

    for (col in corrected_columns) {
        correction_col <- paste0(col, "_correction")
        if (correction_col %in% colnames(corrected_data)) {
            # replace NA in correction with 0 (на всяк випадок)
            corrected_data[[correction_col]][is.na(corrected_data[[correction_col]])] <- 0

            # cumulative correction
            corrected_data[[col]] <- corrected_data[[col]] + cumsum(replace_na(corrected_data[[correction_col]], 0))
        }
    }

    # Remove correction columns
    corrected_data <- corrected_data %>%
        select(-ends_with("_correction"))

    return(corrected_data)
}
```

```{r}
# make both datasets have the same ascending date order

ru_loss_equipment <- ru_loss_equipment %>%
    arrange(date)

# ru_loss_equipment %>%
#     head() %>%
#     select(date, day)

ru_loss_equipment_correction <- ru_loss_equipment_correction %>%
    arrange(date)

ru_loss_equipment_corrected <- apply_corrections(ru_loss_equipment, ru_loss_equipment_correction)

ru_loss_equipment_corrected <- ru_loss_equipment_corrected %>%
    arrange(desc(date))

write.csv(ru_loss_equipment_corrected, "~/Documents/GitHub/scam-project/Project 2/data/2022 Russia Ukraine War/russia_losses_equipment_corrected.csv", row.names = FALSE)
```

Now, we will evaluate the score for each type of equipment!

```{r}
weights <- list(
  aircraft = 69,                 # Aircraft: High cost and strategic value
  helicopter = 50,               # Helicopter: Tactical support and mobility
  tank = 30,                     # Tank: Backbone of ground offensives
  APC = 10,                      # APC: Troop transport
  field.artillery = 15,          # Field Artillery: Combat support
  MRL = 40,                      # MRL: High-impact artillery strikes
  military.auto = 10,            # Military Auto: Logistics and mobility
  fuel.tank = 10,                # Fuel Tank: Operations sustainment
  drone = 20,                    # Drone: Reconnaissance and targeted strikes
  naval.ship = 100,              # Naval Ship: Maritime operations
  anti.aircraft.warfare = 60,    # Anti-Aircraft Warfare: Air defense
  special.equipment = 40,        # Special Equipment: Specialized operational roles
  mobile.SRBM.system = 100,      # Mobile SRBM System: Strategic deterrence
  vehicles.and.fuel.tanks = 30,  # Vehicles and Fuel Tanks: Logistics
  cruise.missiles = 40,          # Cruise Missiles: High-impact strikes
  submarines = 1000,             # Submarines: Strategic naval supremacy
  greatest.losses.direction = 1  # skip this column
)
```

```{r}
ru_loss_equipment_temp <- ru_loss_equipment_corrected %>%
  select(-c(date, day, greatest.losses.direction))

for (equipment in names(weights)) {
  if (equipment %in% colnames(ru_loss_equipment_temp)) {
    ru_loss_equipment_temp[[equipment]] <- ru_loss_equipment_temp[[equipment]] * weights[[equipment]]
  }
}

ru_loss_equipment_temp %>%
  head()

ru_loss_equipment_temp$cdf_loss_score <- rowSums(ru_loss_equipment_temp, na.rm = TRUE)

ru_loss_equipment_evaluated <- cbind(ru_loss_equipment_corrected, cdf_loss_score = ru_loss_equipment_temp$cdf_loss_score)
glimpse(ru_loss_equipment_evaluated)
summary(ru_loss_equipment_temp$cdf_loss_score)
```

Now, we will evaluate the PDF for the equipment losses!

```{r}
ru_loss_equipment_evaluated <- ru_loss_equipment_evaluated %>%
  arrange(date)
```

```{r}
ru_loss_equipment_evaluated_pdf <- ru_loss_equipment_evaluated %>%
  arrange(desc(date)) %>%
  mutate(pdf_loss_score = c(ru_loss_equipment_evaluated$cdf_loss_score[1], diff(ru_loss_equipment_evaluated$cdf_loss_score)))

ru_loss_equipment_evaluated_pdf %>%
  head()
```
```{r}
ru_loss_equipment_evaluated <- ru_loss_equipment_evaluated %>%
  arrange(date)

ru_loss_equipment_evaluated_pdf <- ru_loss_equipment_evaluated_pdf %>%
  arrange(date)

ru_loss_persons <- ru_loss_persons %>%
  arrange(date)
ru_loss_equipment_evaluated <- ru_loss_equipment_evaluated %>%
  arrange(date) %>%
  mutate(equipment_daily_points = cdf_loss_score - lag(cdf_loss_score, default = 0))

head(ru_loss_equipment_evaluated %>% select(date, cdf_loss_score, equipment_daily_points))
ru_loss_persons <- ru_loss_persons %>%
  mutate(personnel = as.numeric(gsub("[^0-9]", "", personnel)))  # Remove non-numeric characters

ru_loss_persons <- ru_loss_persons %>%
  mutate(personnel = replace_na(personnel, 0))

ru_loss_persons <- ru_loss_persons %>%
  arrange(date) %>%
  mutate(personnel_daily_points = personnel - lag(personnel, default = 0))

head(ru_loss_persons %>% select(date, personnel, personnel_daily_points))
```
n


```{r}
ru_loss_equipment_evaluated <- ru_loss_equipment_evaluated %>%
  mutate(date = as.Date(date))

ru_loss_persons <- ru_loss_persons %>%
  mutate(date = as.Date(date))

missiles_daily <- missiles_daily %>%
  mutate(date = as.Date(time_end))

shelling_per_day <- missiles_daily %>%
  group_by(date) %>%
  summarise(shelling_points = sum(launched, na.rm = TRUE))

head(shelling_per_day)
equipment_shelling_merged <- ru_loss_equipment_evaluated %>%
  select(date, equipment_daily_points) %>%
  left_join(shelling_per_day, by = "date")

equipment_shelling_merged <- equipment_shelling_merged %>%
  mutate(shelling_points = replace_na(shelling_points, 0))

head(equipment_shelling_merged)
personnel_shelling_merged <- ru_loss_persons %>%
  select(date, personnel_daily_points) %>%
  left_join(shelling_per_day, by = "date")

personnel_shelling_merged <- personnel_shelling_merged %>%
  mutate(shelling_points = replace_na(shelling_points, 0))

head(personnel_shelling_merged)
```
Model 1: Equipment Loss Points vs. Shelling Points
```{r}
ggplot(equipment_shelling_merged, aes(x = shelling_points, y = equipment_daily_points)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Regression of Daily Equipment Loss Points on Shelling Points",
    x = "Shelling Points (Missiles Launched)",
    y = "Daily Equipment Loss Points"
  ) +
  theme_minimal()
```

```{r}
model1 <- lm(equipment_daily_points ~ shelling_points, data = equipment_shelling_merged)

summary(model1)
```
```{r}
ggplot(personnel_shelling_merged, aes(x = shelling_points, y = personnel_daily_points)) +
  geom_point(alpha = 0.6, color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  labs(
    title = "Regression of Daily Personnel Points on Shelling Points",
    x = "Shelling Points (Missiles Launched)",
    y = "Daily Personnel Points (Killed)"
  ) +
  theme_minimal()
```
	• Intercept (2058.375): This is the estimated equipment loss points when the number of shelling points is zero. While theoretically, shelling points cannot be negative, the intercept provides a baseline level of equipment loss independent of shelling activity.
	•	Slope (24.461): For each additional shelling point (missile launched), the daily equipment loss points increase by approximately 24.46 points. This positive coefficient suggests a direct relationship between shelling activity and equipment losses.
	•	Statistical Significance (Pr(>|t|) < 2e-16): Both the intercept and slope are highly statistically significant, indicating strong evidence against the null hypothesis that these coefficients are zero.
	•	R-squared (0.1309): Approximately 13.09% of the variability in daily equipment loss points is explained by the number of shelling points. While statistically significant, the R-squared value suggests that other factors not included in the model also play a substantial role in explaining equipment losses.
```{r}
model2 <- lm(personnel_daily_points ~ shelling_points, data = personnel_shelling_merged)

summary(model2)
```
	•	Intercept (627.3928): This represents the estimated personnel loss points when the number of shelling points is zero.
	•	Slope (6.3038): For each additional shelling point, the daily personnel loss points increase by approximately 6.30 points. This positive coefficient indicates that increased shelling activity is associated with higher personnel casualties.
	•	Statistical Significance (Pr(>|t|) < 2e-16): Both coefficients are highly significant, providing strong evidence of a relationship between shelling points and personnel casualties.
	•	R-squared (0.22): About 22% of the variability in daily personnel loss points is explained by the number of shelling points. Similar to Model 1, this indicates that while shelling activity is a significant predictor, other factors also influence personnel casualties.
```{r}
par(mfrow = c(2, 2))
plot(model1)

plot(model2)
par(mfrow = c(1, 1))
```

```{r}
massive_shelling_threshold <- 50

equipment_shelling_merged <- equipment_shelling_merged %>%
  mutate(
    massive_shelling = ifelse(shelling_points > massive_shelling_threshold, "Massive", "Non-Massive")
  )

table(equipment_shelling_merged$massive_shelling)

group_means_tech <- equipment_shelling_merged %>%
  group_by(massive_shelling) %>%
  summarise(
    mean_points = mean(equipment_daily_points, na.rm = TRUE),
    sd_points = sd(equipment_daily_points, na.rm = TRUE),
    n = n()
  )

group_means_tech

t_test_tech <- t.test(
  equipment_daily_points ~ massive_shelling,
  data = equipment_shelling_merged
)

t_test_tech

library(ggplot2)

ggplot(equipment_shelling_merged, aes(x = massive_shelling, y = equipment_daily_points)) +
  geom_boxplot(fill = c("lightblue", "salmon")) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4, fill = "white") +
  labs(
    title = "Розподіл очок за техніку між масованими та немасованими обстрілами",
    x = "Тип обстрілу",
    y = "Кількість очок за техніку (daily points)"
  ) +
  theme_minimal()
```